Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-02 Multiresolution_framework/ltx/manuscript.tex
minor edits
added a table for joint state parameter estimation

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace} 
\usepackage[usenames,dvipsnames]{color}

@@

\author[US]{P. Aram\corref{cor1}}
\ead{parham.aram@univ-amu.fr}
\author[DEEE,BEI,UE]{D. R. Freestone}
\ead{dfreestone@bionicear.org} 
\author[UE,CU]{M. Dewar}
\ead{mike.dewar@columbia.edu}
\author[UM]{K. Scerri}
\ead{kenneth.scerri@um.edu.mt}
\author[DEEE,BEI]{D. B. Grayden}
\ead{grayden@unimelb.edu.au}
\author[US]{V. Kadirkamanathan}
\ead{visakan@sheffield.ac.uk}

\address[US]{Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK}
\address[DEEE]{Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia}
\address[BEI]{The Bionic Ear Institute, East Melbourne, VIC, Australia}
\address[UE]{Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK}
\address[CU]{Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA}
\address[UM]{Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta}

@@

\subsection{Stopping rule}
The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}. The stopping rule adopted herein is
\begin{equation}
 \left(\parallel \mathbf{A} \parallel_{F}^{(i)}-\parallel \mathbf{A} \parallel_{F}^{(i-1)}\right)<\epsilon,
 \end{equation}
 where $\epsilon$ is a threshold value and $\parallel \mathbf{A} \parallel_{F}^{(i)}$ and $ \parallel \mathbf{A} \parallel_{F}^{(i-1)}$ are the Frobenius norms  of the successive estimates of $\mathbf{A} $ matrices which is defined by \citet{Meyer2000}
 \begin{equation}
  \parallel \mathbf{A} \parallel_{F}=\sqrt{\sum_{i,j=1}^{n_x}\mid a_{i,j} \mid^2}=\sqrt{\mathrm{tr} (\mathbf A^{\top}\mathbf A)}
 \end{equation}
\subsection{Computational complexity} 
All $n_x^2$ blocks of $\boldsymbol\Gamma_1$ and $\boldsymbol\Gamma_2$  can be computed as a one-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the implementation in \citet{Dewar2009}. The complexity of each step of the algorithm is summarised in Table~\ref{table:MRA-ComputationalComplexity} where for ease of comparison the computational complexity of the method in \citet{Dewar2009} is also identified. It can be seen from the table that the state dimension is an important factor at each iteration of the estimation algorithm, emphasizing the advantage of the proposed algorithm compared to \citet{Dewar2009}. This also suggests the use of an efficient algorithm to choose the minimum number of basis functions (states of the system) while achieving the desired details of the underlying field.  
\begin {table}
\begin{center}
\scalebox{1}{\begin{tabular}{lllc}
\hline \hline
Variable &Equation&Order&Order obtained from \citet{Dewar2009}\\ \hline\\
$\Xi_0, \Xi_1$&\eqref{eq:MRA-Xi0}, \eqref{eq:MRA-Xi1}&$O(Tn_x^2)$& -\\
$\boldsymbol\upsilon_0, \boldsymbol\upsilon_1$&\eqref{eq:epsilon0},\eqref{eq:epsilon1} &$O(n_x^2n_{\theta})$ &$O(n_x^2n_{\theta})+O(n_x^3)$ \\
$\boldsymbol\Upsilon$&\eqref{eq:Epsilon}&$O(n_x^2n_{\theta}^2)$&$O(n_x^4n_{\theta}^2)$\\
$\mathbf A(\boldsymbol\theta)$&\eqref{eq:A_theta}&$O(n_x^3)$&-\\
$\boldsymbol\Lambda_{\theta}$&\eqref{eq:Lambdatheta}&$O(n_x^2n_{\theta})$&-\\
$\hat{\boldsymbol\theta} $&\eqref{eq:MRA-thetahat}&$O(n_{\theta}^3)$&-\\
\hline \hline
\end{tabular}}
 \caption {{\bf The Computational complexity of the estimation algorithm}. A comparison between the algorithm proposed herein and that of \citet{Dewar2009}.} 
\label{table:MRA-ComputationalComplexity}
\end{center}
\end {table}

@@

\renewcommand{\arraystretch}{1.7}
\begin{table*}[!ht]
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Forward initialization} \\ 
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Forward iteration: for $t\in\left\lbrace 0,\cdots,T\right\rbrace $, calculate the predicted state and the predicted covariance matrix } \\
$\hat{\mathbf x}_{t+1}^{f-}=\mathbf A \hat{\mathbf x}_{t}^{f}$  \\ 
$\mathbf P_{t+1}^{f-}=\mathbf A \mathbf P_{t}^{f}\mathbf A^{\top}+\boldsymbol\Sigma_e$  \\
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix} \\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2n_x} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2n_x} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\varepsilon})^{-1}$\\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})$\\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Backward initialization}\\
 $\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Backward iteration: for $t\in\left\lbrace T-1,\cdots,0\right\rbrace $ compute the smoother gain, the smoothed state and the smoothed covariance matrix}\\  
$\mathcal S_{t}=\mathbf P_{t}^{f}\mathbf A^{\top}\left[ \mathbf P_{t +1}^{f-}\right]^{-1}$ \\
 $\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathcal S_t(\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{f-})$ \\
 $\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathcal S_t(\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{f-})\mathcal S_t^\top$\\  
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{5.} Compute the smoothed cross covariance matrix, initialization}\\
$\mathbf M_T=(\mathbf I-\mathcal K_T\mathbf C)\mathbf A\mathbf P_{T-1}^b$\\
\multicolumn{1}{|p{16cm}|}{For $t\in\left\lbrace T-1,\cdots,0\right\rbrace $ compute}\\
$\mathbf M_t= \mathbf P_t^{f}\mathcal S_{t-1}^{\top}+\mathcal S_{t}(\mathbf M_{t+1}-\mathbf A\mathbf P_t^{f} )\mathcal S_{t-1}^{\top}$\\
\hline
\end{tabular}
\caption{\textbf{Algorithm for the RTS smoother} This table shows the steps in the  Rauch-Tung-Striebel smoother algorithm. The lower bound is maximised after each iteration to update the parameter estimates.}
\label{alg:MRA-RTS}
\end{table*}
\renewcommand{\arraystretch}{1} 
\section{Results}\label{sec:MRA-results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using equations \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, allowing a comparison between true and estimated parameters. In doing so, the membrane time constant was set to $\tau = 10$~ms \citep{David2003}, and the sampling time was chosen ten times faster \citep{Stephan2008}, i.e. $T_s = 1$~ms, ensuring the discrepancy between the continuous neural field model and its discrete approximation is small.

@@

\section{Appendix}
\subsection{Inner product of two B-spline scaling functions}\label{ap:InnerProductOfBsplines}
In this section, an analytic formula is derived for the inner product of two B-spline scaling functions. Consider two B-spline scaling functions of order $m$ and $m'$. The support of $N_m\left(s\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e.
\begin{align}\label{eq:APP-SymmetrySplieEq}
 N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right).
\end{align}
A direct consequence of \eqref{eq:APP-SymmetrySplieEq} is 
\begin{align}
 N_{m}\left(r\right)=N_{m}\left(m-r\right).
\end{align}
Therefore
\begin{align}
\int_{-\infty}^{+\infty}N_{m}\left(r-l_{1}\right)N_{m'}\left(r-l_{2}\right)ds&=\int_{-\infty}^{+\infty}N_{m}\left(m-r+l_{1}\right)N_{m'}\left(r-l_{2}\right)dr \nonumber \\
&=\int_{-\infty}^{+\infty}N_{m}\left(m+l_{1}-l_{2}-u\right)N_{m'}\left(u\right)du \nonumber \\
&=\left(N_m \ast N_{m'}\right) \left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m'+l_{2}-l_{1}\right).
\end{align}
\subsection{Scaling function expansion}\label{ap:Induction}  
In this appendix we show by induction that how the scaling function at $j=0$ can be written in terms of the scaling function at level $j$. For $n_1$ in equation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} we have
\begin{align}
 N_{m}\left(r\right)&=\sum_{n_1=0}^{m} \left[\mathbf p\right]_{n_1} N_{m}\left(2r-n_1\right) \label{eq:App-TwoScalepair1},
  \end{align}
which clearly holds from two-scale relation pair (see \eqref{eq:MRA-TwoScalepair1}). Now assuming \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} for $n_{j-1}$ we can write 
\begin{align}\label{eq:app-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_{j-1}=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_{j-1}}N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}}),
\end{align}
in which $N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})$ can be written in terms of $N_m(2^{j}r)$ as
\begin{align}\label{eq:AppendixNj-1intermsofNj}
 &N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})=\nonumber \\
&\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr-2\times k_{n_1,n_2, \dots, n_{j-1}}-n_j)\nonumber \\
&=\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr- k_{n_1,n_2, \dots, n_{j}}).
\end{align}
Substituting \eqref{eq:AppendixNj-1intermsofNj} into \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} completes the proof.

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace} 
\usepackage[usenames,dvipsnames]{color}  
\usepackage{enumerate}

@@

\author[TNG]{P. Aram\corref{cor1}}
\ead{parham.aram@univ-amu.fr}
\author[DEEE,BEI,UE]{D. R. Freestone}
\ead{dfreestone@bionicear.org} 
\author[UE,CU]{M. Dewar}
\ead{mike.dewar@columbia.edu}
\author[UM]{K. Scerri}
\ead{kenneth.scerri@um.edu.mt}
\author[DEEE,BEI]{D. B. Grayden}
\ead{grayden@unimelb.edu.au}
\author[US]{V. Kadirkamanathan}
\ead{visakan@sheffield.ac.uk}

\address[TNG]{Theoretical Neuroscience Group, UMR 1106, Institut de Neurosciences des Systemes, 13385 Marseille, France}
\address[DEEE]{Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia}
\address[BEI]{The Bionic Ear Institute, East Melbourne, VIC, Australia}
\address[UE]{Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK}
\address[CU]{Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA}
\address[UM]{Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta}
 \address[US]{Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK}

@@

\subsection{Stopping rule}
The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}. The stopping rule adopted herein is
\begin{equation}
 \left(\parallel \mathbf{A} \parallel_{F}^{(i)}-\parallel \mathbf{A} \parallel_{F}^{(i-1)}\right)<\epsilon,
 \end{equation}
 where $\epsilon$ is a threshold value and $\parallel \mathbf{A} \parallel_{F}^{(i)}$ and $ \parallel \mathbf{A} \parallel_{F}^{(i-1)}$ are the Frobenius norms  of the successive estimates of $\mathbf{A} $ matrices which is defined by \citet{Meyer2000}
 \begin{equation}
  \parallel \mathbf{A} \parallel_{F}=\sqrt{\sum_{i,j=1}^{n_x}\mid a_{i,j} \mid^2}=\sqrt{\mathrm{tr} (\mathbf A^{\top}\mathbf A)}.
 \end{equation}
\subsection{Computational complexity} 
All $n_x^2$ blocks of $\mathbf U$, $\boldsymbol\Gamma_1$ and $\boldsymbol\Gamma_2$  can be computed as a one-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the implementation in \citet{Dewar2009}. The complexity of each step of the algorithm is summarised in Table~\ref{table:MRA-ComputationalComplexity} where for ease of comparison the computational complexity of the method in \citet{Dewar2009} is also identified. It can be seen from the table that the state dimension is an important factor at each iteration of the estimation algorithm, emphasizing the advantage of the proposed algorithm compared to \citet{Dewar2009}. This also suggests the use of an efficient algorithm to choose the minimum number of basis functions (states of the system) while achieving the desired details of the underlying field.  
\begin {table}
\begin{center}
\scalebox{1}{\begin{tabular}{lllc}
\hline \hline
Variable &Equation&Order&Order obtained from \citet{Dewar2009}\\ \hline\\
$\Xi_0, \Xi_1$&\eqref{eq:MRA-Xi0}, \eqref{eq:MRA-Xi1}&$O(Tn_x^2)$& -\\
$\boldsymbol\upsilon_0, \boldsymbol\upsilon_1$&\eqref{eq:epsilon0},\eqref{eq:epsilon1} &$O(n_x^2n_{\theta})$ &$O(n_x^2n_{\theta})+O(n_x^3)$ \\
$\boldsymbol\Upsilon$&\eqref{eq:Epsilon}&$O(n_x^2n_{\theta}^2)$&$O(n_x^4n_{\theta}^2)$\\
$\mathbf A(\boldsymbol\theta)$&\eqref{eq:A_theta}&$O(n_x^3)$&-\\
$\boldsymbol\Lambda_{\theta}$&\eqref{eq:Lambdatheta}&$O(n_x^2n_{\theta})$&-\\
$\hat{\boldsymbol\theta} $&\eqref{eq:MRA-thetahat}&$O(n_{\theta}^3)$&-\\
\hline \hline
\end{tabular}}
 \caption {{\bf The Computational complexity of the estimation algorithm}. A comparison between the algorithm proposed herein and that of \citet{Dewar2009}.} 
\label{table:MRA-ComputationalComplexity}
\end{center}
\end {table}

@@

\renewcommand{\arraystretch}{1.7}
\begin{table*}[!ht]
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Forward initialization} \\ 
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Forward iteration: for $t\in\left\lbrace 0,\cdots,T\right\rbrace $, calculate the predicted state and the predicted covariance matrix } \\
$\hat{\mathbf x}_{t+1}^{f-}=\mathbf A \hat{\mathbf x}_{t}^{f}$  \\ 
$\mathbf P_{t+1}^{f-}=\mathbf A \mathbf P_{t}^{f}\mathbf A^{\top}+\boldsymbol\Sigma_e$  \\
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix} \\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2n_x} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2n_x} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\varepsilon})^{-1}$\\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})$\\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Backward initialization}\\
 $\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Backward iteration: for $t\in\left\lbrace T-1,\cdots,0\right\rbrace $ compute the smoother gain, the smoothed state and the smoothed covariance matrix}\\  
$\mathcal S_{t}=\mathbf P_{t}^{f}\mathbf A^{\top}\left[ \mathbf P_{t +1}^{f-}\right]^{-1}$ \\
 $\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathcal S_t(\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{f-})$ \\
 $\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathcal S_t(\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{f-})\mathcal S_t^\top$\\  
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{5.} Compute the smoothed cross covariance matrix, initialization}\\
$\mathbf M_T=(\mathbf I-\mathcal K_T\mathbf C)\mathbf A\mathbf P_{T-1}^b$\\
\multicolumn{1}{|p{16cm}|}{For $t\in\left\lbrace T-1,\cdots,0\right\rbrace $ compute}\\
$\mathbf M_t= \mathbf P_t^{f}\mathcal S_{t-1}^{\top}+\mathcal S_{t}(\mathbf M_{t+1}-\mathbf A\mathbf P_t^{f} )\mathcal S_{t-1}^{\top}$\\
\hline
\end{tabular}
\caption{\textbf{Algorithm for the RTS smoother} This table shows the steps in the  Rauch-Tung-Striebel smoother algorithm. The lower bound is maximised after each iteration to update the parameter estimates.}
\label{alg:MRA-RTS}
\end{table*}
\renewcommand{\arraystretch}{1} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\arraystretch}{1.7}
\begin{table*}[!ht]
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Initialization:  } \\ 
\multicolumn{1}{|p{16cm}|}{Choose a bounded random state vector sequence and set $\mathbf P_t$ and $\mathbf M_t$ large} \\
\multicolumn{1}{|p{16cm}|}{Compute} \\
$\mathbf{\Lambda}_{x}=\int_{\Omega}\boldsymbol{\mu}\left(r\right)\boldsymbol{\mu}^\top\left(r\right) \mathrm{d}r$\\
\multicolumn{1}{|p{16cm}|}{Compute all the $n^2_x$ blocks of block matrices $\mathbf U$, $\Gamma_1$ and  $\Gamma_2$ using} \\
$\left[ \mathbf U\right] ^{k,i}=\int_{\boldsymbol \Omega}\left[\boldsymbol\mu(r) \right]_k \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) dr'\right]_{i:} dr$\\
$\left[ \Gamma_1\right]^{j,i} =\sum_{k=1}^{n_x}\left[ \boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1}\right]_{jk} \left[ \mathbf U\right]^{k,i}$\\
$\left[ \Gamma_2\right] ^{j,i}=\sum_{k,m=1}^{n_x}[\mathbf U^{\top}]^{jk} \left[\boldsymbol\Lambda_x^{-1}\boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1} \right]_{km}[\mathbf U]^{mi}$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Estimate the connectivity kernel parameters} \\
$\boldsymbol\Xi_0=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right)$\\
$\boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right)$\\
$\boldsymbol\upsilon_0=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Gamma_1]_{j,i}$\\
$\boldsymbol\upsilon_1=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_1]^{j,i}$\\
$\boldsymbol\Upsilon=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_2]^{j,i}$\\
$\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Form a state-space model}\\
\multicolumn{1}{|p{16cm}|}{Compute all the elements of $\boldsymbol\Lambda_{\theta}$ using} \\
$\left[ \boldsymbol\Lambda_{\theta}\right] _{k,i}=\left[ \mathbf U\right]^{k,i}\boldsymbol\theta$\\
\multicolumn{1}{|p{16cm}|}{Compute transition matrix $\mathbf A(\boldsymbol\theta)$ } \\
$\mathbf A(\boldsymbol \theta) = T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I$\\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Compute the smoothed state $\mathbf x_t^b$, the smoothed covariance,$\mathbf P^b_t$ and cross-covariance $\mathbf M_t$ for $t\in\left\lbrace 0,\cdots,T\right\rbrace $ using the RTS smoother given in Table~\ref{alg:MRA-RTS}.}\\  
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{5.} Convergence: go back to Step~2 unless }\\
$ \left(\parallel \mathbf{A} \parallel_{F}^{(i)}-\parallel \mathbf{A} \parallel_{F}^{(i-1)}\right)<\epsilon$\\
\hline
\end{tabular}
\caption{\textbf{State and parameter estimation for the MRAIDE model.} This table shows the steps for joint state and parameter estimation using the EM based algorithm.}
\label{alg:EMsteps}
\end{table*}
\renewcommand{\arraystretch}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:MRA-results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using equations \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, allowing a comparison between true and estimated parameters. In doing so, the membrane time constant was set to $\tau = 10$~ms \citep{David2003}, and the sampling time was chosen ten times faster \citep{Stephan2008}, i.e. $T_s = 1$~ms, ensuring the discrepancy between the continuous neural field model and its discrete approximation is small.

@@

\section{Appendix}
\subsection{Inner product of two B-spline scaling functions}\label{ap:InnerProductOfBsplines}
In this section, an analytic formula is derived for the inner product of two B-spline scaling functions. Consider two B-spline scaling functions of order $m$ and $m'$. The support of $N_m\left(s\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e.
\begin{align}\label{eq:APP-SymmetrySplieEq}
 N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right).
\end{align}
A direct consequence of \eqref{eq:APP-SymmetrySplieEq} is 
\begin{align}
 N_{m}\left(r\right)=N_{m}\left(m-r\right).
\end{align}
Therefore
\begin{align}
\int_{-\infty}^{+\infty}N_{m}\left(r-l_{1}\right)N_{m'}\left(r-l_{2}\right)dr&=\int_{-\infty}^{+\infty}N_{m}\left(m-r+l_{1}\right)N_{m'}\left(r-l_{2}\right)dr \nonumber \\
&=\int_{-\infty}^{+\infty}N_{m}\left(m+l_{1}-l_{2}-u\right)N_{m'}\left(u\right)du \nonumber \\
&=\left(N_m \ast N_{m'}\right) \left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m'+l_{2}-l_{1}\right).
\end{align}
\subsection{Scaling function expansion}\label{ap:Induction}  
In this appendix we show by induction that how the scaling function at $j=0$ can be written in terms of the scaling function at level $j$. For $n_1$ in equation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} we have
\begin{align}
 N_{m}\left(r\right)&=\sum_{n_1=0}^{m} \left[\mathbf p\right]_{n_1} N_{m}\left(2r-n_1\right) \label{eq:App-TwoScalepair1},
  \end{align}
which clearly holds from two-scale relation pair (see \eqref{eq:MRA-TwoScalepair1}). Now assuming \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} for $n_{j-1}$ we can write 
\begin{align}\label{eq:app-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_{j-1}=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_{j-1}}N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}}),
\end{align}
in which $N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})$ can be written in terms of $N_m(2^{j}r)$ as
\begin{align}\label{eq:AppendixNj-1intermsofNj}
 &N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})=\nonumber \\
&\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr-2\times k_{n_1,n_2, \dots, n_{j-1}}-n_j)\nonumber \\
&=\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr- k_{n_1,n_2, \dots, n_{j}}).
\end{align}
Substituting \eqref{eq:AppendixNj-1intermsofNj} into \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} completes the proof.
